### 2014-11-02 fundamentals/impl.tex

\Cyclus, on the other hand, is primarily written in \texttt{C++} and relies on 
libraries supported by Linux and UNIX (including Ubuntu and OSX) platforms. 
Furthermore, the core infrastructure and related archetypes are free and 
open source, BSD-3-clause licensed. No part of \Cyclus is proprietary or based 
on \gls{COTS}, Windows-based software. It can be easily deployed 
on large computer systems, such as \acrlong{HTC} systems.

Cyclopts (\TODO{cite}), a proof of principle design and implementation of such a
system, uses UW-Madison's HTCondor HTC infrastructure to perform sensitivity
studies on \Cyclus' resource exchange optimization solvers. To date, it has been
used run over $10^4$ jobs, or (\TODO{compute time}) total compute hours, using
the \Cyclus library via its resource exchange \gls{API}.
The HTC infrastructure has also been utilized to run and collect
information from full \Cyclus simulations running in parallel on $10^3$
machines reliably for order $10^5$ simulations.

@@

In particular, since the clean plugin architecture loads libraries without any
modifications to the \Cyclus kernel, closed-source archetypes can be used with
the simulator alongside open-source archetypes. This architecture
allows closed-source libraries (e.g., those representing sensitive nuclear
processes and subject to export control) to be developed and licensed privately.

>>>>

\Cyclus, on the other hand, is primarily written in \texttt{C++} and relies on 
libraries supported by Linux and UNIX (including Ubuntu and OSX) platforms. 
Furthermore, the core infrastructure and related archetypes are free and 
open source, BSD-3-clause licensed. No part of \Cyclus is proprietary or based 
on \gls{COTS}, Windows-based software. It can be easily deployed 
on large computer systems, such as \gls{HTC} systems.

Cyclopts (\TODO{cite}), a proof of principle design and implementation of such a
system, uses UW-Madison's HTCondor \gls{HTC} infrastructure to perform sensitivity
studies on \Cyclus' resource exchange optimization solvers. To date, it has been
used run over $10^4$ jobs, or (\TODO{compute time}) total compute hours, using
the \Cyclus library via its resource exchange \gls{API}.
The \gls{HTC} infrastructure has also been utilized to run and collect
information from full \Cyclus simulations running in parallel on $10^3$
machines reliably for order $10^5$ simulations.

@@

In particular, since the clean plug-in architecture loads libraries without any
modifications to the \Cyclus kernel, closed-source archetypes can be used with
the simulator alongside open source archetypes. This architecture
allows closed-source libraries (e.g., those representing sensitive nuclear
processes and subject to export control) to be developed and licensed privately.

### 2014-11-02 fundamentals/qa.tex

Multiple strategies collectively known as \emph{\gls{QA}} have 
been invented over the years to mitigate the structural and algorithmic errors
on the part of simulators themselves. These include \emph{\gls{VV}}
\cite{boehm_software_1989}, \emph{\gls{UQ}} 
\cite{sacks_design_1989}, testing, and others. \gls{VV} and \gls{UQ} derive from 
a distinctly computational science background while testing \emph{et al.}
proceed from a software development bent.

Nuclear enginerering code quality is often governed by NQA-1, an ASME specification 
whose latest revision appeared in 2009 \cite{asme_nqa-1a-2009_2009}. This is primarily 
used for designing reactors. However, it is general enough to apply to 
\Cyclus. \Cyclus has adopted an \emph{agile} development process 
\cite{larman_agile_2004}, 
interpreting NQA-1 in a manner similar to NEAMS \cite{neams_nuclear_2013} or 
PyNE \cite{biondo_quality_2014}.

\Cyclus acknowledges that quality assurance is an on-going process throughout the 
entire life of the code. As a simulator where most modeling descisions are made 
by third party archetype developers or users, the most important feature of
\gls{QA} 
is verification. Validation may be impossible and \gls{UQ} is beyond the scope.

Verification may be defined as the question, ``Is \Cyclus being built correctly?'' 
This question can be answered by the software development of notions of testing,
documentation, version control, style guidelines, and continuous integration. 
This suite of process controls supplies mechanisms for reliable and reproducible 
software. The impetus to implement these correctly is even stronger in a scientific 
context because of the emphasis on reproducibility and provenance. These are 
discussed below individually.

Validation on the other hand may be defined as the question, 
``Is \Cyclus the correct tool?''
Since \Cyclus is alone in its class as an agent-based fuel cycle simulator longitudnal 
validation is not possible. Still code-to-code comparisons with fuel cycle
simulators with other modeling paradigms are underway, if nascent. However, such 
exercises are more likely to bring into relief the differences between the modeling
paradigms than be useful for QA and validation.

>>>>

Multiple strategies collectively known as \emph{\gls{QA}} have 
been invented over the years to mitigate the structural and algorithmic errors
on the part of simulators themselves. These include \emph{verification and validation}
(V\&V) \cite{boehm_software_1989}, \emph{uncertainty quantification} (UQ) 
\cite{sacks_design_1989}, testing, and others. V\&V and UQ come from 
a distinctly computational science background while testing \emph{et al.} come from 
a software development bent.

Nuclear enginerering code quality is often governed by NQA-1, an ASME specification 
whose latest revision appeared in 2009 \cite{asme_nqa-1a-2009_2009}. This is primarily 
used for designing reactors. However, it is general enough to apply to 
cyclus. Cyclus has adopted an \emph{agile} development process 
\cite{larman_agile_2004}, 
interpreting NQA-1 in a manner similar to NEAMS \cite{neams_nuclear_2013} or 
PyNE \cite{biondo_quality_2014}.

Cyclus acknowledges that quality assurance is an on-going process throughout the 
entire life of the code. As a simulator where most modeling descisions are made 
by third party archetype developers or users, the most important feature of QA 
is verification. Validation may be impossible and UQ is beyond the scope.

Verification may be defined as the question, ``Is cyclus being built correctly?'' 
To answer this question we turn to the software development of notions of testing,
documentation, version control, style guidelines, and continuous integration. 
This suite of process controls supplies mechanims for reliable and reproducible 
software. The impetus to implement these correctly is even stronger in a scientific 
context because of the emphasis on reproducibility and provenance. These are 
discussed below individually.

Validation on the other hand may be defined as the question, 
``Is cyclus the correct tool?''
Since cyclus is alone in its class as an agent-based fuel cycle simulator logitudnal 
validation is not possible. Still code-to-code comparisons with fuel cycle
simulators with other modeling paradigms are underway, if nascent. However, such 
exercises are more likely to bring into relief the differences between the modeling
paradigms than be useful for QA and validation.
